On the presentation of orthogonal groups over polynomial rings

### 2019-08-01 dl-paper.tex
Proving 5.30

\section{Proof of the first part of the injectivity theorem}
We start this section by recalling basic notation and facts pertaining to hermitian K-theory.
First of all, recall that the stable orthogonal K-groups $\KO_i(R) = \KO_i(\infty, R)$ are precisely the hermitian K-groups ${}_{-1}\!L_i(R)$ introduced by Karoubi in~\cite{Ka80}. We will need yet another redesignation for these groups $\KO_i(R)$, which comes from the theory of {\it higher Grothendieck--Witt groups of schemes}.
Recall that this theory, developed by M.~Schlichting, is a modern broad generalization of the classical hermitian K-theory of rings developed in~\cite{Ka80}.
We refer the reader to~\cite[\S~2]{FRS12} and~\cite[\S~2]{AF17} for the introduction to it.

@@

We will also need the following property of relative Steinberg groups which is a special case of {\it ``Tulenbayev lifting property''} discussed in~\cite{LS17}.
\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$,
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}[ur] & \St(\Phi, R) } \]
Then there exists a diagonal arrow which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.
 In the general case the assertion of the lemma is a corollary of amalgamation theorem~\cite[Theorem~9]{S15} 
  (see~\cite[\S~2]{LS17} for more details).
\end{proof}

@@

\subsection{Construction of a model for $\St(\Phi, M[X] + A[X^{-1}])$}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
Let $A$ be a local ring with maximal ideal $M$.
We denote by $B$ the ring $M[X] + A[X^{-1}]$ considered as a graded subring of $A[X, X^{-1}]$. The notation of the present subsection is consistent with~\cref{sec:presentation} if we set $t = X^{-1}$, $I = M$.

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X^{-1}]) &  } \] 
Denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, A[X, X\inv])$.
Denote by $V$ the subset of \[\St(\Phi, A[X, X^{-1}]) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$. This action is well defined since $G^0_M \subset \overline{G}^{\geq 0}_M$. Whenever two triples $v_1, v_2 \in V$ lie in the same orbit we use the notation $v_1 \sim v_2$ and say that they are congruent. The same notation $h_1 \cong h_2$ is also used for the elements of $\St(\Phi, A[X\inv])$ that are right-adjacent with respect to $G^0_M$.

We denote by $\overline{V}$ the set of orbits of this action.
We use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ mapping each orbit $(x, [h], u)$ to the coset $p(x, h, u)G_M^0$.

We start by defining for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for the triples $(g, h, u)$ satisfying $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ belongs to $P_\alpha(m)$ for some $m \in M$, then we define $t_\alpha(aX^{-1})$ by the following identity:
\begin{equation} \label{T_1} t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\end{equation}

@@

This action satisfies the following additional property. If $v = (g, [h], u) \in \overline{V}$ is such that $\overline{p}(v) = \overline{1} = G_M^0$ then for any $g_1 \in \St(\Phi, A[X\inv])$ holds $\overline{p}(j_-(g_1) \cdot v) = \overline{1}$.
\end{prop}
\begin{proof}
  By~\cref{superfluous-relations} and Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1} the action of $\St^{\leq 1}(\Phi, B)$ on $\overline{V}$ given by~\eqref{T_1} and~\eqref{T_leq0} is well-defined. On the other hand, the group $\St^{\leq 1}(\Phi, B)$ is isomorphic to $\St(\Phi, B)$ by~\cref{lemma33} which implies the first assertion.

@@

We need to verify the following property: for $g_1 \in \overline{G}_M^{\geq 0}$ and $v\in \overline{V}$ one has $\overline{p}(t(g_1) \cdot v) = g_1 \cdot \overline{p}(v)$. It suffices to verify this for $g = z_\alpha(f, \xi)$, $f \in M[X]$, $\xi\in A[X]$.

>>>>

\section{Proof of the first part of the injectivity theorem} \label{firstPart}
We start this section by recalling basic notation and facts pertaining to hermitian K-theory.
First of all, recall that the stable orthogonal K-groups $\KO_i(R) = \KO_i(\infty, R)$ are precisely the hermitian K-groups ${}_{-1}\!L_i(R)$ introduced by Karoubi in~\cite{Ka80}. We will need yet another redesignation for these groups $\KO_i(R)$, which comes from the theory of {\it higher Grothendieck--Witt groups of schemes}.
Recall that this theory, developed by M.~Schlichting, is a modern broad generalization of the classical hermitian K-theory of rings developed in~\cite{Ka80}.
We refer the reader to~\cite[\S~2]{FRS12} and~\cite[\S~2]{AF17} for the introduction to it.

@@

We will also need the following property of relative Steinberg groups which is a special case of {\it ``Tulenbayev lifting property''} discussed in~\cite[\S~2]{LS17}.
\begin{lemma}\label{lem:lemma32} Let $\Phi$ be a simply-laced root system of rank $\geq 3$,
Consider the following commutative square of canonical maps.
\[ \xymatrix{
    \St(\Phi, B, I) \ar[r] \ar[d] & \St(\Phi, B) \ar[d] \\
    \St(\Phi, R, I) \ar[r] \ar@{-->}^t[ur] & \St(\Phi, R) } \]
Then there exists a diagonal map $t$ which makes the diagram commute.   
\end{lemma} 
\begin{proof}
 Notice that $R$ is isomorphic to the principal localisation of $B$ at $X$
  and that $I$ is uniquely $X$-divisible in the sense of~\cite[\S~4]{LS17}.
 Thus, in the special case $\Phi = \rA_3$ the assertion of the lemma follows from~\cite[Theorem~3]{LS17}.
 In the general case the assertion of the lemma is a corollary of amalgamation theorem~\cite[Theorem~9]{S15} 
  (see~\cite[\S~2]{LS17} for more details).
\end{proof}

@@

\subsection{Construction of a model for $\St(\Phi, A[X\inv] + M[X])$}
Throughout this subsection $\Phi$ denotes arbitrary irreducible simply-laced root system of rank $\geq 3$, unless stated otherwise.
Let $A$ be a local ring with maximal ideal $M$.
As in~\cref{firstPart} we denote by $B$ the ring $A[X\inv] + M[X]$ considered as a graded subring of $A[X, X\inv]$. The notation of the present subsection is consistent with~\cref{sec:presentation} if we set $t = X\inv$, $I = M$.

For shortness we denote the subgroup $\overline{\St}(\Phi, A, M)$ by $G_M^0$.
Notice that $G_M^0$ can be considered as a subgroup of $\St(\Phi, A[X, X\inv])$.
\[ \xymatrix{ G_M^0 \ar@{^{(}->}[r]^{i_+} \ar@{^{(}->}[d]^{i_-} & \overline{\St}(\Phi, A[X], M[X]) \ar^{j_+^M}[d] \ar@{^{(}->}[r] & \St(\Phi, A[X]) \ar[dl]^{j_+} \\
              \St(\Phi, A[X\inv]) \ar[r]^{j_-} & \St(\Phi, A[X, X\inv]) &  } \] 
We denote by $\overline{G}^{\geq 0}_M$ the image of the map $j_+^M$. 
Denote by $V$ the subset of \[\St(\Phi, A[X, X\inv]) \times \St(\Phi, A[X\inv]) \times (1 + M)^\times\] consisting of those triples $(g, h, u)$ for which
\[p(g, h, u) := g \cdot j_-(h) \cdot \{ X, u \}\] belongs to $\overline{G}_M^{\geq 0}$.

We let $h_0 \in G_M^0$ act on $V$ on the right by $(g, h, u) \cdot h_0 = (g, h \cdot i_-(h_0), u)$. This action is well defined since $G^0_M \subset \overline{G}^{\geq 0}_M$. Whenever two triples $v_1, v_2 \in V$ lie in the same orbit we use the notation $v_1 \sim v_2$ and say that they are {\it congruent}. The same notation $h_1 \sim h_2$ is also used for the elements of $\St(\Phi, A[X\inv])$ lying in the same left $G^0_M$-coset.

We denote by $\overline{V}$ the set of orbits of the action of $G_M^0$.
We use the notation $(g, [h], u)$ for the elements of $\overline{V}$.
We denote by $\overline{p}$ the function $\overline{V} \to \overline{G}^{\geq 0}_M/G_M^0$ sending each $(g, [h], u) \in V$ to the left coset $p(g, h, u)G_M^0$.

We start by defining for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for the triples $(g, h, u)$ satisfying $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u)$ belongs to $P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is defined via the following identity:
\begin{equation} \label{T_1} t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\end{equation}

@@

This action satisfies the following additional property. If $v = (g, [h], u) \in \overline{V}$ is such that $\overline{p}(v) = \overline{1} = G_M^0$ then for any $h_1 \in \St(\Phi, A[X\inv])$ holds $j_-(h_1) \cdot v = (h_1 \cdot g, [h \cdot h_1^{-1}], u)$ and, moreover, $\overline{p}(j_-(h_1) \cdot v) = \overline{1}$.
\end{prop}
\begin{proof}
  By~\cref{superfluous-relations} and Lemmas~\ref{R3_leq0_1}--\ref{R2_0_1} the action of $\St^{\leq 1}(\Phi, B)$ on $\overline{V}$ given by~\eqref{T_1} and~\eqref{T_leq0} is well-defined. On the other hand, the group $\St^{\leq 1}(\Phi, B)$ is isomorphic to $\St(\Phi, B)$ by~\cref{lemma33} which implies the first assertion.

@@

\begin{lemma}
For $g_1 \in \overline{G}_M^{\geq 0}$ and $v\in \overline{V}$ one has $\overline{p}(t(g_1) \cdot v) = g_1 \cdot \overline{p}(v)$, where $t$ is the map from~\cref{lem:lemma32}.
\end{lemma}
\begin{proof}
 It suffices to verify this property for the generators $g = z_\alpha(f, \xi)$ of the subgroup $\overline{G}^{\geq 0}_M$ ($f \in M[X]$, $\xi\in A[X]$).
 %Notice that in the case $\xi = 0$ the assertion immediately follows from~\eqref{T_leq0} since $t(x_\alpha(f)) = x_\alpha(f)$. 
 We verify the assertion by induction on the degree of $\xi$ in $X$.

 First of all, consider the case $\xi = aX$, $a\in A$. Notice that
 \[t(z_\alpha(f, aX)) = [x_{\beta}(X^{-1}) \cdot x_{\beta-\alpha}(a),\ x_{\alpha-\beta}(-fX) \cdot x_{-\beta}(afX^2)].\]
 \[ [x_\beta(X^{-1})^{x_{-\alpha}(aX)},\ x_{\alpha-\beta}(Xf)^{x_{-\alpha}(aX)}] \]
\end{proof}

\begin{multline}z_{\alpha+\beta}(f\xi, X) = x_\alpha(\epsilon f) \cdot x_{-\beta}(-fX) \cdot x_{\beta}(fX\xi^2) \cdot x_{\alpha+\beta}(f \xi) \cdot \\ \cdot z_\alpha(-\epsilon f, -\epsilon X\xi) \cdot
  x_{-\alpha}(-\epsilon fX^2\xi^2) \cdot x_{-\alpha-\beta}(- f X^2 \xi) \cdot z_{-\beta}(fX, -\xi)\end{multline}

\begin{multline}
 z_{\alpha+\beta}(f, ax) = x_\alpha(\epsilon Xf) x_{-\beta}(-faX^2) \cdot x_{\beta}(af) \cdot x_{\alpha+\beta}(f) \cdot \\ \cdot z_\alpha(-\epsilon Xf, -\epsilon a) \cdot x_{-\alpha}(-\epsilon a^2fX) \cdot x_{-\alpha-\beta}(- f a^2X^2) \cdot z_{-\beta}(afX^2, -X^{-1})
\end{multline}

$s := fX$, $\eta := X^{-1}$, $\xi = aX$
